Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation

### 2021-06-06 manuscript/manuscript.tex
minor title fix

\title{Natural Selection is Unlikely to Explain Why Species Get a Thin Slice of $\pi$}
%\title{Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation}

@@

% TODO recheck node height and other p-values

@@

A major impediment in quantifying Lewontin's Paradox has been estimating census
population sizes across many taxa, especially for extremely abundant,
cosmopolitan species that define the upper limit of ranges. Previous work has
surveyed the literature for census size estimates
\parencite{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size,
or qualitative categories as proxies for census size
\parencite{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
between genomic estimates of diversity and census population sizes, I first
approximate census population sizes for 172 metazoan taxa (Figure
\ref{fig:figure-nc}). My approach predicts population densities from body sizes
using a previously-observed linear relationship that holds across metazoans
(Figure 1-figure supplement 1 \ref{suppfig:damuth};
\cite{Damuth1981-st,Damuth1987-sg}). Then, from geographic occurrence data, I
estimate range sizes.  Finally, I estimate population size as the product of
these predicted densities and range estimates (see Methods:
\Newnameref{sec:methods-popsize}). Note that the relationship between
population density and body size is driven by energy budgets, and thus reflects
macroecological equilibria \parencite{Damuth1987-sg}. Consequently, population
sizes are underestimated for taxa like humans and their domesticated species,
and overestimated for species with anthropogenically reduced densities or
fragmented ranges. For example, the population size of \emph{Lynx lynx} is
likely around 50,000 \parencite{Iucn2020-ap} which is around two orders of
magnitude smaller than my estimate. Additionally, the range size estimates do
not consider whether an area has unsuitable habitat, and thus may be
overestimated for species with particular niches or patchy habitats. While
these methods to estimate census size are crude and approximate, they can be
efficiently calculated for numerous taxa and are sufficient to estimate the
scale of Lewontin's Paradox (see Supplementary Information
\Newnameref{supp:popsize-valid} for more on validation based on biomass and
other approaches).

@@

Notably, this relationship has few outliers and is relatively homoscedastic.
This is in part because of the log-log scale, in contrast to previous work
\parencite{Nei1984-zi,Soule1976-he}; see Figure 1-supplement figure 2 for a
version on a log-linear scale. However, it is noteworthy that few taxa have
diversity estimates below $10^{-3.5}$ differences per basepair. Those that do,
lynx (\emph{Lynx Lynx}), wolverine (\emph{Gulo gulo}), and Massasauga
rattlesnake (\emph{Sistrurus catenatus}) face habitat fragmentation and
declining population sizes. These three species are all in the IUCN Red List,
but are listed as least concern (though their presence in the Red List
indicates they are of conservation interest). In Supplementary Information
Section \Newnameref{supinfo:div-iucn}, I explore more about the relationships
between IUCN Red List status, diversity, and population size.

@@

\section{Figure Supplement}

\setcounter{figure}{0}
\renewcommand{\thefigure}{}
\renewcommand{\figurename}{Figure 1-figure supplement 1}

>>>>

% \title{Natural Selection is Unlikely to Explain Why Species Get a Thin Slice of $\pi$}
\title{Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation}

@@

@@

A major impediment in quantifying Lewontin's Paradox has been estimating census
population sizes across many taxa, especially for extremely abundant,
cosmopolitan species that define the upper limit of ranges. Previous work has
surveyed the literature for census size estimates
\parencite{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size,
or qualitative categories as proxies for census size
\parencite{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
between genomic estimates of diversity and census population sizes, I first
approximate census population sizes for 172 metazoan taxa (Figure
\ref{fig:figure-nc}). My approach predicts population densities from body sizes
using a previously-observed linear relationship that holds across metazoans
(see Supplementary Figure \ref{suppfig:damuth};
\cite{Damuth1981-st,Damuth1987-sg}). Then, from geographic occurrence data, I
estimate range sizes.  Finally, I estimate population size as the product of
these predicted densities and range estimates (see Methods:
\Newnameref{sec:methods-popsize}). Note that the relationship between
population density and body size is driven by energy budgets, and thus reflects
macroecological equilibria \parencite{Damuth1987-sg}. Consequently, population
sizes are underestimated for taxa like humans and their domesticated species,
and overestimated for species with anthropogenically reduced densities or
fragmented ranges. For example, the population size of \emph{Lynx lynx} is
likely around 50,000 \parencite{Iucn2020-ap} which is around two orders of
magnitude smaller than my estimate. Additionally, the range size estimates do
not consider whether an area has unsuitable habitat, and thus may be
overestimated for species with particular niches or patchy habitats. While
these methods to estimate census size are crude and approximate, they can be
efficiently calculated for numerous taxa and are sufficient to estimate the
scale of Lewontin's Paradox (see Supplementary Information
\Newnameref{supp:popsize-valid} for more on validation based on biomass and
other approaches).

@@

Notably, this relationship has few outliers and is relatively homoscedastic.
This is in part because of the log-log scale, in contrast to previous work
\parencite{Nei1984-zi,Soule1976-he}; see Supplementary Figure
\ref{suppfig:figure-1-linear} for a version on a log-linear scale. However, it
is noteworthy that few taxa have diversity estimates below $10^{-3.5}$
differences per basepair. Those that do, lynx (\emph{Lynx Lynx}), wolverine
(\emph{Gulo gulo}), and Massasauga rattlesnake (\emph{Sistrurus catenatus})
face habitat fragmentation and declining population sizes. These three species
are all in the IUCN Red List, but are listed as least concern (though their
presence in the Red List indicates they are of conservation interest). In
Supplementary Information Section \Newnameref{supinfo:div-iucn}, I explore more
about the relationships between IUCN Red List status, diversity, and population
size.

@@
